---
title: "DAG for anemia in FinOCC"
author: "Mikko Arvas"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DiagrammeR)
library(ggdag)
```


# DAGs

# Overall DAG

```{r}
#based on https://www.erikigelstrom.com/articles/causal-graphs-in-r-with-diagrammer/

grViz("
digraph {
  graph [ranksep = 0.2]

  node [shape = plaintext]
    A    [label = 'Anemia', shape = box]
    I    [label = 'Iron stores']

    HC  [label = 'Hormonal contraception', shape = box ]
    Mens [label = 'Menstruation' , shape = box]
    Pre  [label = 'Pregnancy' , shape = box]
    Cb  [label = 'Child birth' , shape = box]
    MS  [label = 'Metabolic syndrome' , shape = box]
    D   [label = 'Diet']
    IS  [label = 'Iron supplements']
    M   [label = 'Migraine', shape = box]
    Ag  [label = 'Age' , shape = box]
    B   [label = 'Blood donation', shape = box]
    O   [label = 'Overweight', shape = box ]
    S   [label = 'Smoking' , shape = box]
    Ac  [label = 'Acne' , shape = box]
    An  [label = 'Anorexia' , shape = box]
    PE  [label = 'Physical excercise']
    CI  [label = 'Other diagnoses?' , shape = box]
    C   [label = 'Cancer' shape = box]
    
    
    #SEP  [label = 'Socioeconomic position', shape = box]

  edge [minlen = 2]
    I->A
    Mens->I
    HC->Mens #HC will reduce menstrual flow
    Mens -> HC # If you have excessive menstrual flow you are more likely to be recommended HC
    Mens -> Pre
    HC->Pre
    Cb -> HC
    Pre->I
    Pre->Cb
    Cb->I
    D->I
    IS->I
    M->HC
    MS->A # Overweight (and smoking?) cause IDA 
    MS -> HC # 'The risk of ischemic stroke in patients using combined oral contraceptives is increased in patients with additional stroke risk factors, including smoking, hypertension, and migraine with aura'

    Ag -> HC
  #  O -> A
    O -> MS
    B -> I
    S -> MS
    S -> HC
    Ac -> HC
    D -> An
    An -> Mens
    PE -> Mens
    PE -> I
    C -> A
    HC -> A

{ 
  rank = same; I; A
#  rank = same; CI; MS
  }
}
")
```
Which diagnoses before anemia diagnoses are bases for exclusion?

  -Everything we use as basis of exclusion, we don't have to worry about.

What else could impact on the choice of hormonal contraception?

Should "Diet" have a direct arrow to menstruation?

Should "Age" have a direct arrow to menstruation?

Should we separate combined oral contraceptive pill and  progesterone only analysis?

# DAG of what we have in FinOCC

```{r}

anemia_dag <- ggdag::dagify(
  HC ~  Mens, # If you have excessive menstrual flow you are more likely to be recommended HC
  Mens ~ HC, # If you have HC you won't have excessive menstrual flow
  Mens ~ An , # Anorexia can stop excessive menstrual flow
  A ~ Mens , # Excessive menstrual flow can cause anemia
  Pre ~ HC, # HC will block pregnancy
  HC ~ Pre, # Pregnancy can affect your use/choise of HC
  HC ~ Cb , # Cb can affect your use/choise of HC
  Cb ~ Pre, #Pregnancy can cause child birth
  HC ~ M , #Migraine can effect choise of HC
  A ~ MS  , # Obesity can cause IDA 
  HC ~ MS, # 'The risk of ischemic stroke in patients using combined oral contraceptives is increased in patients with additional stroke risk factors, including smoking, hypertension, and migraine with aura'
  HC ~ Ag, # Age can affect your choise of HC
  Pre ~ Ag, # Age reduces your likely hood of getting pregnant
  MS ~ O  , # Obesity causes metabolic syndrome
  HC ~ Ac , # Acne can affect your choise of HC

  A ~ C, # cancer can cause anemia
  A ~ HC, # hormonal contraception can protect from anemia, but should this go only through the menstruation variable?
  labels = c(
    "A" = 'Anemia',
    "HC" = 'Use/Choise of HC',
    "Mens"= 'Ex. mens. flow' ,
    "Pre" = 'Pregnancy' ,
    "Cb" = 'Child birth' ,
    "MS" = 'Metabolic syndrome' ,
    "M" =  'Migraine',
    "Ag" = 'Age' ,
    "O"  = 'Obesity',
    "Ac"  = 'Acne' ,
    "An" = 'Anorexia',
    "C" = 'Cancer'
  ),
  exposure = "HC",
  outcome = "A"
)

ggdag::ggdag(anemia_dag, # the dag object we created
             text = FALSE, # this means the original names won't be shown
             use_labels = "label") #+ # instead use the new names
```

```{r}
ggdag_paths(anemia_dag)
```

```{r}
ggdag_adjustment_set(anemia_dag)
```


Excessive menstrual flow will cause a loop that needs to be blocked. How?
  -Exclude all persons with excessive menstrual flow diagnosis
  -Similarly cancer could cause all kinds of loops so all persons with cancer diagnosis are best excluded
  
  

```{r}

anemia_01_dag <- ggdag::dagify(
  #HC ~  Mens, # If you have excessive menstrual flow you are more likely to be recommended HC
  #Mens ~ HC, # If you have HC you won't have excessive menstrual flow
  Pre ~ HC, # HC will block pregnancy
  HC ~ Pre, # Pregnancy can affect your use/choise of HC
  HC ~ Cb , # Cb can affect your use/choise of HC
  Cb ~ Pre, #Pregnancy can cause child birth
  HC ~ M , #Migraine can effect choise of HC
  A ~ MS  , # Obesity can cause IDA 
  #A ~ Mens , # Excessive menstrual flow can cause anemia
  HC ~ MS, # 'The risk of ischemic stroke in patients using combined oral contraceptives is increased in patients with additional stroke risk factors, including smoking, hypertension, and migraine with aura'
  HC ~ Ag, # Age can affect your choise of HC
  Pre ~ Ag, # Age reduces your likely hood of getting pregnant
  MS ~ O  , # Obesity causes metabolic syndrome
  HC ~ Ac , # Acne can affect your choise of HC
#  Mens ~ An , # Anorexia can stop excessive menstrual flow
#  A ~ C,
  A ~ HC,
  labels = c(
    "A" = 'Anemia',
    "HC" = 'Use/Choise of HC',
    "Mens"= 'Ex. mens. flow' ,
    "Pre" = 'Pregnancy' ,
    "Cb" = 'Child birth' ,
    "MS" = 'Metabolic syndrome' ,
    "M" =  'Migraine',
    "Ag" = 'Age' ,
    "O"  = 'Obesity',
    "Ac"  = 'Acne' ,
    "An" = 'Anorexia',
    "C" = 'Cancer'
  ),
  exposure = "HC",
  outcome = "A"
)

ggdag::ggdag(anemia_dag, # the dag object we created
             text = FALSE, # this means the original names won't be shown
             use_labels = "label") #+ # instead use the new names
```

```{r}
ggdag_paths(anemia_01_dag)
```

```{r}
ggdag_adjustment_set(anemia_01_dag)
```
Hence, the model would be:

A ~  hormonal_contraception + metabolic_syndrome

