---
title: "DAG for anemia in FinOCC"
author: "Mikko Arvas"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DiagrammeR)
library(ggdag)
library(causaleffect)
library(igraph)
library(tidyverse)
library(dagitty)
```

TODO:


-what do we do with age and time ? Can we break loops with some timing?

  
  -if we consider only women with out kids does age matter for inflammatory status, choise of HC or something else ?
  
  
  - Eeva said
  
    -younger are more likely to get combination pills, they can used to treat acne and menstrual disorders and combination pill raise risk of venous thromboembolism
  
    -for progestine pills age should not matter
    
    -> so there is age affect but also possibly the age effect depends on the hc type
    
  -how about how long a product has been used? the longer you have used it the more time it has had to have an effect? and this also is dependent on age. Could Muriel give me some pointers on how to address it? Could it be a percentage of age or is an interaction better as the estimate really don't matter? Do we have the dates of when the products were ordered? Or should we do a survival analysis?
  
-should we add a propensity score analysis with a score on how likely one is to use HC

-trial emulation?

# Functions

```{r}
from_ggdag_2_diagrammer <- function(ggdag.object) {
  tmp <- tidy_dagitty(ggdag.object)$data
 #tmp <- tidy_dagitty(anemia_dag)$data
  edges <- tmp %>% 
    select(name, to) %>% 
    rename("from"="name") %>% 
    na.omit()
  
   nodes <- tmp %>% 
   select(name,label) %>%
   #  select(name) %>% 
   rename("id"="name") %>%
   distinct() %>%
   na.omit()
  g <- create_graph(
    directed = TRUE,
    attr_theme = "tb"
  )
  g<- add_nodes_from_table(g,
                           table = nodes,
                           label_col = label)
  g <- add_edges_from_table(graph =g,
                            table=edges,
                            from_col = from,
                            to_col = to,
                            from_to_map=id_external)
  invisible(g)
}

```


```{r}
from_ggdag_2_igraph <- function(dagify.object) {
  tmp <- tidy_dagitty(dagify.object)$data
  tmp <- tmp %>% select(name, to) %>% 
  as.matrix() %>% # igraph wants a matrix
  na.omit() #there was A -> NA , maybe to spesify outcome
  tmp <- graph_from_edgelist(tmp, directed = TRUE)
  invisible(tmp)
}

```


# Pastricha DAG with hormonal contraception

Based on "Iron deficiency" 
Sant-Rayn Pasricha, Jason Tye-Din, Martina U Muckenthaler, Dorine W Swinkels 2021 Lancet
http://dx.doi.org/10.1016/S0140-6736(20)32594-0

```{r}

pastricha_dag <- ggdag::dagify(
  II ~ VD, # Low haem iron content (eg, from a vegetarian or vegan diet)
  II ~ ConInh, # Concomitant consumption of inhibitors of iron absorption (eg, calcium or tea)
  II ~ InAcid, # Inadequate acidification
  II ~ InMyc, # Intestinal mycosal dysfunction
  IS ~ II, #When iron intake is inadequate to meet requirements or to compensate for physiological or pathological losses, body iron stores become depleted
  A ~ Iinfl, #In patients with inflammation, withholding of iron from the plasma promotes iron deficient erythropoiesis and anaemia despite adequate body iron stores (functional iron deficiency). Obesity
  A ~ IS, # Absolute ID occurs when iron stores are insufficient to meet the needs of the individual, and is particularly common in young children (younger than 5 years) and premenopausal (especially pregnant) women
  IS ~ Mens,  # The main causes of absolute ID are excessive blood loss...
  Mens ~ HC,
  IS ~ InLoss,  # Gastrointestinal blood loss is the most important cause of ID in men and postmenopausal women
  IS ~ BD,
  IS ~ Cb,
  Cb ~ Pre,
  
  
  
  labels = c(
    "A" = 'Iron\ndeficiency\nanemia',
    "HC" = 'Use/Choise\nof HC',
    "Mens"= 'Menstruation' ,
    "Pre" = 'Pregnancy' ,
    "Cb" = 'Child\nbirth' ,
    "II" = 'Iron\nintake',
    "IS" = 'Iron\nstores',
   "Iinfl" = 'Inflammation',
   "InLoss" = 'Gastrointestinal\nblood loss',
   "Erosion" = "Erosion",
   "CC" = 'Colon cancer',
   "InflBo" = 'Inflamatory bowel disease',
   "BD" = 'Blood\ndonation',
   "ConInh" = 'Consumption\nof inhibitors',
   "InMyc" = 'Intestinal\ndysfunction',
   "InAcid" = 'Inadequate\nacidification',
   "VD" = 'Low\niron\ndiet'
   
  ),
  exposure = "HC",
  outcome = "A"
)

pastricha_dag.diag <- from_ggdag_2_diagrammer(pastricha_dag)
render_graph(
  pastricha_dag.diag, 
#  layout="tree"
)
```

```{r}
latents(pastricha_dag) <- c("II","IS")
exposures(pastricha_dag) <- c("HC")
outcomes(pastricha_dag) <- c("A")

```


```{r}
ggdag_adjustment_set(pastricha_dag)
#how to define the latent variables to get the corract adjustement set
```

‘Backdoor Paths Unconditionally Closed’ means that, assuming the DAG we have drawn is correct, we may obtain an unbiased estimate of X on Y without including additional variables.

```{r}
ggdag_instrumental(pastricha_dag)
```

# Pastricha DAG with added complexity


```{r}
pastricha_dag_mens <- ggdag::dagify(
  II ~ VD, # Low haem iron content (eg, from a vegetarian or vegan diet)
  II ~ ConInh, # Concomitant consumption of inhibitors of iron absorption (eg, calcium or tea)
  II ~ InAcid, # Inadequate acidification
  II ~ InMyc, # Intestinal mycosal dysfunction
  IS ~ II, #When iron intake is inadequate to meet requirements or to compensate for physiological or pathological losses, body iron stores become depleted
  A ~ Iinfl, #In patients with inflammation, withholding of iron from the plasma promotes iron deficient erythropoiesis and anaemia despite adequate body iron stores (functional iron deficiency). Obesity
  A ~ IS, # Absolute ID occurs when iron stores are insufficient to meet the needs of the individual, and is particularly common in young children (younger than 5 years) and premenopausal (especially pregnant) women
  IS ~ Mens,  # The main causes of absolute ID are excessive blood loss...
  Pre ~ IS , # if you are low own iron you are less likely to get pregnant
  Mens ~ HC,
  IS ~ InLoss,  # Gastrointestinal blood loss is the most important cause of ID in men and postmenopausal women
  IS ~ BD, # Blood donation can cause iron deficiency
  BD ~ IS, # If you are low own iron you can't get to donate 
  IS ~ Cb,
  IS ~ Pre,
  Cb ~ Pre,
  Mens ~ IS, # Your iron store can influence your menstruation
  HC ~ Mens, # Your menstruation can influece you choise of HC
  HC ~ Iinfl, # obesity and mentabolic disease can influence your choise of HC
  HC ~ Cb, # having kids will impact on you choice of hormonal contraception
  HC ~ Age, # age affect your choice of contraception, should there HC X age interaction
  Iinfl ~ Age, # the older you are the more inflamed you get, "spline"?
  A ~ Age, # due to iron requirements of growth younger women tend to be more iron deficient aas they have not yet requperated from growth. Should this be a "spline" ?   
  
  
  labels = c(
    "A" = 'Iron\ndeficiency\nanemia',
    "HC" = 'Use/Choise\nof HC',
    "Mens"= 'Menstruation' ,
    "Pre" = 'Pregnancy' ,
    "Cb" = 'Child\nbirth' ,
    "II" = 'Iron\nintake',
    "IS" = 'Iron\nstores',
   "Iinfl" = 'Inflammation',
   "InLoss" = 'Gastrointestinal\nblood loss',
   "Erosion" = "Erosion",
   "CC" = 'Colon cancer',
   "InflBo" = 'Inflamatory bowel disease',
   "BD" = 'Blood\ndonation',
   "ConInh" = 'Consumption\nof inhibitors',
   "InMyc" = 'Intestinal\ndysfunction',
   "InAcid" = 'Inadequate\nacidification',
   "VD" = 'Low\niron\ndiet',
   "Age" = 'Age'
   
  ),
  exposure = "HC",
  outcome = "A"
)

pastricha_dag.diag <- from_ggdag_2_diagrammer(pastricha_dag_mens)
render_graph(
  pastricha_dag.diag, 
#  layout="tree"
)
```


```{r}
latents(pastricha_dag_mens) <- c("II","IS")
exposures(pastricha_dag_mens) <- c("HC")
outcomes(pastricha_dag_mens) <- c("A")

```


```{r}
ggdag_adjustment_set(pastricha_dag_mens)
#how to define the latent variables to get the corract adjustement set
```




```{r}
ggdag_instrumental(pastricha_dag_mens)
```



# Pastricha DAG with added complexity and exclusions

Child birth and pregnancy create backdoor paths. Exclude all data prior to first pregnancy.
Use of hormonal contraception to treat excessive menstrual flow. Exclude persons with excessive menstrual flow.
Similarly cancer could cause all kinds of loops so all persons with cancer diagnosis are best excluded.

```{r}
pastricha_dag_mens_no <- ggdag::dagify(
  II ~ VD, # Low haem iron content (eg, from a vegetarian or vegan diet)
  II ~ ConInh, # Concomitant consumption of inhibitors of iron absorption (eg, calcium or tea)
  II ~ InAcid, # Inadequate acidification
  II ~ InMyc, # Intestinal mycosal dysfunction
  IS ~ II, #When iron intake is inadequate to meet requirements or to compensate for physiological or pathological losses, body iron stores become depleted
  A ~ Iinfl, #In patients with inflammation, withholding of iron from the plasma promotes iron deficient erythropoiesis and anaemia despite adequate body iron stores (functional iron deficiency). Obesity
  A ~ IS, # Absolute ID occurs when iron stores are insufficient to meet the needs of the individual, and is particularly common in young children (younger than 5 years) and premenopausal (especially pregnant) women
  IS ~ Mens,  # The main causes of absolute ID are excessive blood loss...
  Mens ~ HC,
  IS ~ InLoss,  # Gastrointestinal blood loss is the most important cause of ID in men and postmenopausal women
  IS ~ BD, # Blood donation can cause iron deficiency
  BD ~ IS, # If you are low own iron you can't get to donate 
  Mens ~ IS, # Your iron store can influence your menstruation
  HC ~ Iinfl, # obesity and mentabolic disease can influence your choise of HC
  HC ~ Age, # age affect your choice of contraception, should there HC X age interaction
  Iinfl ~ Age, # the older you are the more inflamed you get, "spline"?
  A ~ Age, # due to iron requirements of growth younger women tend to be more iron deficient aas they have not yet requperated from growth. Should this be a "spline" ? 
 
  labels = c(
    "A" = 'Iron\ndeficiency\nanemia',
    "HC" = 'Use/Choise\nof HC',
    "Mens"= 'Menstruation' ,
    "Pre" = 'Pregnancy' ,
    "Cb" = 'Child\nbirth' ,
    "II" = 'Iron\nintake',
    "IS" = 'Iron\nstores',
   "Iinfl" = 'Inflammation',
   "InLoss" = 'Gastrointestinal\nblood loss',
   "Erosion" = "Erosion",
   "CC" = 'Colon cancer',
   "InflBo" = 'Inflamatory bowel disease',
   "BD" = 'Blood\ndonation',
   "ConInh" = 'Consumption\nof inhibitors',
   "InMyc" = 'Intestinal\ndysfunction',
   "InAcid" = 'Inadequate\nacidification',
   "VD" = 'Low\niron\ndiet',
   "Age" = 'Age'
   
  ),
  exposure = "HC",
  outcome = "A"
)

pastricha_dag.diag <- from_ggdag_2_diagrammer(pastricha_dag_mens_no)
render_graph(
  pastricha_dag.diag, 
#  layout="tree"
)
```


```{r}
latents(pastricha_dag_mens_no) <- c("II","IS")
exposures(pastricha_dag_mens_no) <- c("HC")
outcomes(pastricha_dag_mens_no) <- c("A")

```


```{r}
ggdag_adjustment_set(pastricha_dag_mens_no)
```


```{r}
ggdag_instrumental(pastricha_dag_mens_no)
```
# Previous experiments

test
