---
title: "EmergencyProcess_Flowchart_InfusionOnTheScene"
author: "Mikko Arvas"
date: '2022-06-19'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DiagrammeR)
```


```{r}
#based on https://www.erikigelstrom.com/articles/causal-graphs-in-r-with-diagrammer/

grViz("
digraph {
  graph [ranksep = 0.2]

  node [shape = plaintext]
    A    [label = 'Anemia', shape = box]
    I    [label = 'Iron stores']

    HC  [label = 'Hormonal contraception', shape = box ]
    Mens [label = 'Menstruation' , shape = box]
    Pre  [label = 'Pregnancy' , shape = box]
    Cb  [label = 'Child birth' , shape = box]
    MS  [label = 'Metabolic syndrome' , shape = box]
    D   [label = 'Diet']
    IS  [label = 'Iron supplements']
    M   [label = 'Migraine', shape = box]
    Ag  [label = 'Age' , shape = box]
    B   [label = 'Blood donation', shape = box]
    O   [label = 'Overweight', shape = box ]
    S   [label = 'Smoking' , shape = box]
    Ac  [label = 'Acne' , shape = box]
    An  [label = 'Anorexia' , shape = box]
    PE  [label = 'Physical excercise']
    CI  [label = 'Other diagnoses?' , shape = box]
    C   [label = 'Cancer' shape = box]
    
    
    #SEP  [label = 'Socioeconomic position', shape = box]

  edge [minlen = 2]
    I->A
    Mens->I
    HC->Mens
    Mens -> Pre
    HC->Pre
    Cb -> HC
    Pre->I
    Pre->Cb
    Cb->I
    D->I
    IS->I
    M->HC
    MS->A
    Ag -> HC
  #  O -> A
    O -> MS
    B -> I
    S -> MS
    Ac -> HC
    An -> D
    An -> Mens
    PE -> Mens
    PE -> I
    C -> A
{ 
  rank = same; I; A
#  rank = same; CI; MS
  }
}
")
```
Which diagnoses before anemia diagnoses are bases for exclusion?
What else could impact on the choice of hormonal contraception?
Should "Diet" have a direct arrow to menstruation?
Should "Age" have a direct arrow to menstruation?

-Paljon vuotavalle ei kuparikierukkaa
-Tukosriski  -> ei yhdistelmä
-Tupakointi -> tukosriski
-käypähoito suositukset
-yhdistelmä ehkäisy vai progestiini eri kaaviot?


